---
title: "Data Visualization on Heart Disease Data Set using R"
author: Chih-Hua Chang
date: "2022-09-30"
output: pdf_document
---
Data Set:
https://archive.ics.uci.edu/ml/datasets/Heart+Disease (processed.cleveland.data)

```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
processed.cleveland = read.csv("/Users/fionachang/Desktop/data607/hw4/processed.cleveland.data",
header=FALSE)
```

Replace value based on condition
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
processed.cleveland$V14[processed.cleveland$V14 != 0] <- 1
```


# Count Plot
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
library(ggplot2)
library(tidyverse)
#continuous value → categorical value
processed.cleveland$V2 <- as.factor(processed.cleveland$V2)

count_df <- processed.cleveland %>% group_by(V2) %>% count(V2)

#Count plot
ggplot(count_df, aes(x = V2, y = n)) +
	geom_col(position = "dodge") + #group bar plot
	theme(plot.title = element_text(size=12)) + #white background, title size
	labs(title="Count by Sex", y="Count",x="Sex") + #change x, y axis title
	scale_x_discrete(labels=c('female', 'male')) + #change x axis labels
	geom_text(aes(label = n), 
						colour = "white", size = 3, 
						vjust = 1.5, position = position_dodge(.9)) #bar labels
```
- Total number of male: 206, Total number of female: 97
- The total number of male is twice larger the total number of female


# Group Count Plot
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
#count by group
count_df <- processed.cleveland %>% group_by(V2) %>% count(V14)

#continuous value → categorical value
count_df$V2 <- as.factor(count_df$V2)
count_df$V14 <- as.factor(count_df$V14)

#group count plot
ggplot(count_df, aes(x = V2, y = n, fill = V14)) +
	geom_col(position = "dodge") + #group bar plot
	theme(plot.title = element_text(size=12)) + #white background, title size
	labs(title="Count by Sex with Presence/Absence of Heart Disease",  y="Count",x="Sex")+ 
  #change x, y axis title
	guides(fill=guide_legend(title="Heart Disease")) + #change legend title
	scale_x_discrete(labels=c('female', 'male')) + #change x axis labels
	scale_fill_discrete(labels=c('absence', 'presence')) + #change legend labels
	geom_text(aes(label = n), colour = "white", size = 3, vjust = 1.5, 
	          position = position_dodge(.9)) #bar labels
```


# Pie Plot
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
processed.cleveland$V2 <- as.factor(processed.cleveland$V2)
processed.cleveland$V14 <- as.factor(processed.cleveland$V14)

count_df <- processed.cleveland %>% group_by(V2) %>% count(V14) %>% ungroup(V14) %>% 
                                    mutate(perc = `n` / sum(`n`)) %>% 
                                    arrange(perc) %>% 
                                    mutate(labels = scales::percent(perc))

ggplot(data=count_df, aes(x=" ", y=n, group=V14,  fill=V14)) +
	geom_bar(width = 1, stat = "identity") +
	geom_text(aes(label = labels),position =position_stack(vjust = 0.5)) +
	coord_polar("y", start=0) +
	facet_grid(.~ V2) + 
	theme_void() + 
	ggtitle("Sex") +
	guides(fill=guide_legend(title="Heart Disease")) + 
	scale_fill_discrete(labels = c("absence", "presence"))
```
- The percentage of having heart disease in male is larger than female.

```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
ggplot(processed.cleveland, aes(x = V1)) +
	geom_bar() +
	labs(title="Age Distribution", y="Count",x="Age")
```

```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
count_df <- processed.cleveland %>% group_by(V1) %>% count(V14)
ggplot(count_df, aes(x = V1, y = n, fill = V14)) +        
	geom_col(position = "dodge") + #group bar plot         
	theme(plot.title = element_text(size=12)) +         
	labs(title="Age Distribution with Presence/Absence of HD ", y="Count",x="Age") + #change x, y axis title         
	guides(fill=guide_legend(title="Heart Disease")) + #change legend title         
	scale_fill_discrete(labels=c('absence', 'presence')) #change legend labels
```

- In the age group of 30-55, The salmon bar is higher than the blue bar, which shows that there are less cases in heart disease
- From age 55 to age 65, there are more presence of heart disease than the absence of heart disease.
- And after age 65, it seems that the presence and absence of heart disease is equal
- So, by the graph we assume that age around 55 to 65 are more likely to have heart disease.


# Violin Plot

http://www.sthda.com/english/wiki/ggplot2-violin-plot-quick-start-guide-r-software-and-data-visualization

```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
#continuous value → categorical value
processed.cleveland$V2 <- as.factor(processed.cleveland$V2) 

ggplot(processed.cleveland, aes(x=V2, y=V1, fill=V14)) +
	geom_violin(trim=FALSE) + 
	theme(plot.title = element_text(size=12)) +
	labs(title="Age by Sex with Presence/Absence of Heart Disease", y="Age",x="Sex") +
	scale_x_discrete(labels=c('female', 'male')) + guides(fill=guide_legend(title="Heart Disease")) + 
	scale_fill_discrete(labels=c('absence', 'presence'))
```

- The shape of female and male in presence of heart disease is very different.
- The shape of female is more extreme, the age range of having heart disease are around 55-65, and there is much less case out of the age range
- However, the age range of having heart disease in male is more prevalent,  though there are still a relatively larger area in age 60’s
- The relationship of age and the presence of heart disease is more sensitive in female than in male.



# Scatter Plot
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
ggplot(processed.cleveland, aes(V5,V8,color = V14)) + 
	geom_point()+
	labs(y="Maximum Heart Rate Achieved ",x="Serum Cholestoral in mg/dl ") + 
	guides(color = guide_legend(title = "Heart Disease")) +  #legend title
	scale_colour_discrete(labels=c('absence', 'presence'))  #lengend labels```
```


# Logistic Regression
```{r, echo=TRUE, message = FALSE, out.width="50%",out.height="50%", fig.align = 'center'}
#sample training data 
train <- sample(303, 242)

#train logistic regression
logRegDef<-glm(V14 ~ V1 + V2 + V3 + V4 + V5 + V6 + V7 + V8 + V9 + V10 + V11 
											+ V12 + V13,              
	family=binomial, data=processed.cleveland, subset=train)

#training result summary
summary(logRegDef)$coefficients
summary(logRegDef)

#test logistic regression
logRegDef.predict<-predict(logRegDef,
	newdata=processed.cleveland[-train,], type="response")

#y predict by sigmoid
ypred<-ifelse(logRegDef.predict<1/2, 0, 1)

#y values
table(processed.cleveland$V14[-train])

#accuracy
mean(ypred == processed.cleveland[-train,]$V14)  
sum(ypred!=processed.cleveland$V14[-train])/(303-242) 
```


